\speculake: Hiding Malware in Speculative Execution

### 2018-12-11 paper/related-work.tex
add related work to blog

Spectre builds upon prior work on cache side channels, and similarly uses them
to leak information from
processes~\cite{percival2005cache,zhang2012cross,osvik2006cache}. In \speculake,
we use the branch predictor as a \emph{covert channel}~\cite{lampson1973note}
between the trigger program and malware payload, allowing the malware's
(speculative) execution path to be influenced by the trigger.

@@

\speculake builds on Spectre~\cite{spectre} and Meltdown~\cite{meltdown} which
leverage speculative execution to leak sensitive information from vulnerable
processes.  Follow up work has identified several new Spectre variants,
including speculative buffer overflows and speculative store
bypass~\cite{kiriansky2018speculative,spec-store-bypass}, and has investigated
additional ways to leak information using branch
predictors as a side channel~\cite{evtyushkin2018branchscope}.
Researchers have also leveraged Spectre to demonstrate web-based
vulnerabilities~\cite{genkin2018drive,schwarz2018netspectre} as well as to leak
keys and other information from the hardware isolation provided by Intel
SGX~\cite{spectre_sgx,chen2018sgxpectre,foreshadow}.
\speculake likewise takes advantage of speculative execution, but with the goal
of hiding arbitrary computation from reverse engineering, rather than extracting
secrets from vulnerable programs. \speculake additionally benefits from
new Spectre variants: as we showed, speculative buffer overflows (``Spectre~1.1'')
can be used as an alternative trigger for malware.

>>>>

%programs~\cite{evtyushkin2016jump, lee2017inferring, aciiccmez2007predicting}
Spectre builds upon prior work on cache side channels, and similarly uses them
to leak information from
processes~\cite{percival2005cache,zhang2012cross,osvik2006cache}. In \speculake,
we use the branch predictor as a \emph{covert channel}~\cite{lampson1973note}
between the trigger program and malware payload, allowing the malware's
(speculative) execution path to be influenced by the trigger.

@@

\speculake builds on Spectre~\cite{spectre} and Meltdown~\cite{meltdown} which
leverage speculative execution to leak sensitive information from vulnerable
processes.  Follow up work has identified several new Spectre variants,
including speculative buffer overflows and speculative store
bypass~\cite{kiriansky2018speculative,spec-store-bypass}, and has investigated
additional ways to leak information using branch
predictors as a side channel~\cite{evtyushkin2018branchscope}.
Researchers have also leveraged Spectre and speculative execution more generally
to demonstrate web-based
vulnerabilities~\cite{genkin2018drive,schwarz2018netspectre} as well as to leak
control flow, keys, and other information from the hardware isolation provided by Intel
SGX~\cite{spectre_sgx,chen2018sgxpectre,foreshadow,lee2017inferring}.
Spectre has additionally been proposed as a way to thwart taint tracking by using speculative
execution to copy data between buffers~\cite{necessaryBlog}.
\speculake likewise takes advantage of speculative execution, but with the goal
of hiding arbitrary computation from reverse engineering, rather than extracting
secrets from vulnerable programs. \speculake also benefits from
new Spectre variants: as we showed, speculative buffer overflows (``Spectre~1.1'')
can be used as an alternative trigger for malware.
